hbm_paper_results

### 2024-02-05 paper_lbcc_results.Rmd
figure 4

abs_delta_summary.df <- cent_deltas_df %>%
  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta") %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(study) %>%
  summarise(study_mean_abs.delta=mean(abs.delta),
            study_mean_delta=mean(delta),
            mean_age=mean(age_days),
            study_max_abs.delta=max(abs.delta),
            study_n=n())
```

@@

delta_sum.df_piv <- delta_sum.df %>%
  pivot_wider(id_cols=c(study, study_n, mean_age, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))

@@

delta_sum.df %>%
  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age)) +
  geom_count(aes(size=sex_n)) +
  geom_smooth(method="lm") +
  facet_wrap(~sex)

delta_sum.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")
```

@@

```{r}
delta_sum.df %>%
  ggplot(aes(x=mean_age, y=mean_abs.delta, color=sex)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")

delta_sum.df_piv %>%
  ggplot(aes(x=mean_age, y=(study_mean_abs.delta*100))) +
  geom_count(aes(size=study_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("grDevices::Oslo", name = "Proportion \nFemale") +
  geom_smooth(method="lm")
```

```{r}
# age
summary(lm(study_mean_abs.delta ~ mean_age, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ mean_age, data=delta_sum.df_piv))
```

are studies w younger subjects changing more because there are fewer of them?
check for association between study size and age
```{r}
summary(lm(mean_age ~ study_n, data=delta_sum.df_piv))
fit <- lm((study_mean_abs.delta*100) ~ mean_age + study_n, data=delta_sum.df_piv)
summary(fit)
sjPlot::plot_model(fit, type = "pred", terms = "mean_age")

summary(lm((study_mean_abs.delta*100) ~ mean_age + study_n + sex_prop_Female, data=delta_sum.df_piv))
```

@@

```{r}
# %female
summary(lm(study_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ sex_prop_Female, data=delta_sum.df_piv))
```

```{r}
# age*%female
summary(lm(study_mean_abs.delta ~ mean_age*sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ mean_age*sex_prop_Female, data=delta_sum.df_piv))
```

@@

```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="age_days") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```
```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="sexMale:age_days") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```

@@

```{r}
deltas_df %>%
  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm")

deltas_df %>%
  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
  geom_histogram(aes(y=after_stat(c(
      count[group==1]/sum(count[group==1]),
      count[group==2]/sum(count[group==2])
    )*100)), alpha=0.4, position="identity", bins=60) +
  ylab("Percentage")

deltas_df %>%
  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth(method="lm")

deltas_df %>%
  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm")
```

@@

**Figure out what features didn't converge & remove them from brain atlas** - seems to be happening automatically...?

RERUN MODELS WITH AGE IN YEARS!!!

```{r}
term_names <- c(sexMale = "Sex", age_days = "Age (Days)", 'sexMale:age_days' = "Sex*Age (Days)")

#Regional
regional.delta_plt <- cent_delta_lms.df_toplt %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ estimate*100)) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = estimate)) +
  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))

#Global
global.delta_plt <- cent_delta_lms.df_toplt %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ estimate*100),
         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA))) %>%
  dplyr::select(!any_of(c("label", "region", "hemi"))) %>% #drop cols corresponding to dk regions
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             side = c("coronal"),
             aes(fill = estimate)) +
  facet_grid(term ~ pheno_cat) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
```

>>>>

abs_delta_summary.df <- cent_deltas_df %>%
  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta") %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(study) %>%
  summarise(study_mean_abs.delta=mean(abs.delta),
            study_mean_delta=mean(delta),
            mean_age=mean(age_days),
            mean_age_yrs=mean(age_days)/365.25,
            study_max_abs.delta=max(abs.delta),
            study_n=n())
```

@@

delta_sum.df_piv <- delta_sum.df %>%
  pivot_wider(id_cols=c(study, study_n, mean_age, mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))

@@

delta_sum.df %>%
  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age_yrs)) +
  geom_count(aes(size=sex_n)) +
  geom_smooth(method="lm") +
  facet_wrap(~sex)

delta_sum.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age_yrs)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")
```

```{r}
# age
summary(lm(study_mean_abs.delta ~ mean_age_yrs, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ mean_age_yrs, data=delta_sum.df_piv))
```

@@

```{r}
delta_sum.df %>%
  ggplot(aes(x=mean_age_yrs, y=mean_abs.delta, color=sex)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")

age_effects.plt <- delta_sum.df_piv %>%
  ggplot(aes(x=mean_age_yrs, y=(study_mean_abs.delta*100))) +
  geom_count(aes(size=study_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("grDevices::Oslo", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#060B14FF") +
  theme_linedraw(base_size = 11) +
  labs(size="Study N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")
```

are studies w younger subjects changing more because there are fewer of them?
check for association between study size and age
```{r}
summary(lm(mean_age_yrs ~ study_n, data=delta_sum.df_piv))
fit <- lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n, data=delta_sum.df_piv)
summary(fit)
sjPlot::plot_model(fit, type = "pred", terms = "mean_age_yrs")

summary(lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_piv))
```

@@

delta ~ % female

```{r}
# %female
summary(lm(study_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ sex_prop_Female + study_n, data=delta_sum.df_piv)) #add study size covariate
summary(lm(study_mean_delta ~ sex_prop_Female + study_n + mean_age, data=delta_sum.df_piv)) #add study size & age covariates
```

```{r}
delta_sum.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100), color=study, fill=study)) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=study))

sex_effects.plt <- delta_sum.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=(study_mean_delta*100))) +
  geom_count(aes(size=study_n, color=mean_age_yrs)) +
  geom_smooth(method="lm", color="#006027FF") +
  paletteer::scale_color_paletteer_c("grDevices::Greens 2", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="Study N", x = "Proportion Female", y = "Mean Difference in Centile Scores")

delta_sum.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100))) +
  geom_boxplot()
```

```{r}
# age*%female
summary(lm(study_mean_abs.delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
```

@@

```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="age_yrs") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```

```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="sexMale:age_yrs") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```

@@

```{r}
cent_deltas_df %>%
  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm")

cent_deltas_df %>%
  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
  geom_histogram(aes(y=after_stat(c(
      count[group==1]/sum(count[group==1]),
      count[group==2]/sum(count[group==2])
    )*100)), alpha=0.4, position="identity", bins=60) +
  ylab("Percentage")

cent_deltas_df %>%
  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth(method="lm")

cent_deltas_df %>%
  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm")
```

@@

```{r}
term_names <- c(sexMale = "Sex", age_yrs = "Age", 'sexMale:age_yrs' = "Sex*Age")

#Regional
regional.delta_plt <- cent_delta_lms.df_toplt %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ estimate*100)) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = estimate)) +
  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))

#Global
global.delta_plt <- cent_delta_lms.df_toplt %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ estimate*100),
         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA))) %>%
  dplyr::select(!any_of(c("label", "region", "hemi"))) %>% #drop cols corresponding to dk regions
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             side = c("coronal"),
             aes(fill = estimate)) +
  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names)), switch = 'y') + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
```

#FIGURE 4

```{r}
# study demographics
fig4 <- ggarrange(
  ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          nrow=1, ncol=2, 
          align = "h", 
          common.legend = TRUE,
          legend="bottom",
          widths=c(2, 1)),

# featurewise effects
ggarrange(global.delta_plt, regional.delta_plt,
          nrow=1, ncol=2, 
          align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(1, 2.5)
          ),
nrow=2, ncol=1,
heights=c(1.5, 1),
labels=c("A", "C")
),
# mean effects
ggarrange(age_effects.plt,
sex_effects.plt,
nrow=2, ncol=1,
align="hv",
labels=c("B", "D")),
common.legend=FALSE,
nrow=1, ncol=2,
widths=c(1, 1.3)
)

fig4_annotate <- annotate_figure(fig4, top = text_grob("Figure 4. Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
fig4_annotate
ggsave("figs/figure_4.jpeg", fig4_annotate, width=12, height=10, units="in")
```
